Analysis of Cartesian Genetic Programming's Evolutionary Mechanisms

### 2014-03-26 CGP Journal/main.tex
Condensing to fit onto 15 pages.

In total this creates 9 algorithm combinations.  Of these, our previous work
has been reported showing the effectiveness of the duplication methods with \emph{Normal} ordering~\cite{goldman:2013:cgpwaste}
as well as analysis of all three ordering methods with \emph{Single}~\cite{goldman:2013:ordering}.
This means there remain 4 novel combinations to be tested.
This is also the first time the ordering methods
have been rigorously tested for quality, as much of the work in~\cite{goldman:2013:ordering}
was analysis oriented.

@@

Somewhat unintuitively, node reactivation and active node semantic changes appear
to be highly correlated.  This is apparent in the fact that the \textbf{No Reactivation}
column in each table is almost perfectly sorted, even though rows were sorted using the
\textbf{Active Unchanged} column.  Again the number of active nodes seems the likely
cause.  Having fewer active nodes reduces the number of introns, making it more
difficult for inactive nodes to become active.  Fewer active nodes also decreases
how many inactive nodes could have ever been active.

Potentially the most interesting results come from the control combination of
\emph{Skip-Normal}.  As problem difficulty increases, \emph{Skip-Normal} produces
an increasingly large number of offspring with no change in active node behavior,
with little to no node reactivation, and a high chance of changing nodes before
they are reactivated.  This suggests two important features: \emph{Skip} is likely
saving a significant number of evaluations, and \emph{Normal} CGP is likely using
inactive nodes primarily to inject random behavior.  The former comes from the
argument that a large part of the active unchanged individuals are also likely
actively identical to their parents.  The latter is supported by the fact that
\emph{Skip-Normal} so infrequently reactivates a node with tested behavior.  More
often than not offspring connecting in inactive nodes are testing that node's behavior
for the first time.

>>>>

In total this creates 9 algorithm combinations.  Of these, our previous work
%TODO Changed here
has shown the effectiveness of the duplication methods with \emph{Normal} ordering~\cite{goldman:2013:cgpwaste}
as well as analysis of all three ordering methods with \emph{Single}~\cite{goldman:2013:ordering}.
This means there remain 4 novel combinations to be tested.
This is also the first time the ordering methods
have been rigorously tested for quality, as much of the work in~\cite{goldman:2013:ordering}
was analysis oriented.

@@

Somewhat unintuitively, node reactivation and active node semantic changes appear
to be highly correlated.  This is apparent in the fact that the \textbf{No Reactivation}
%TODO Changed here
column in each table is almost perfectly sorted, even though rows were sorted by the
\textbf{Active Unchanged} column.  Again the number of active nodes is the likely
cause.  Having fewer active nodes reduces the number of introns, making it more
difficult for inactive nodes to become active.  Fewer active nodes also decreases
how many inactive nodes could have ever been active.

Potentially the most interesting results come from the control combination of
\emph{Skip-Normal}.  As problem difficulty increases, \emph{Skip-Normal} produces
an increasingly large number of offspring with no change in active node behavior,
with little node reactivation, and a high chance of changing nodes before
they are reactivated.  This suggests two important features: \emph{Skip} is likely
saving a significant number of evaluations, and \emph{Normal} CGP is likely using
inactive nodes primarily to inject random behavior.  The former comes from the
%TODO Changed here
argument that most of the active unchanged individuals are also likely
actively identical to their parents.  The latter is supported by the fact that
\emph{Skip-Normal} infrequently reactivates a node with tested behavior with
offspring frequently connecting in untested inactive nodes.
